---
title: "Visualizations"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## Bird Migration
Goal: to investigate if a higher number of birds are hit during their migration period through comparing the relationship between the months of migration (March to April, August to November), the quantity of birds, and the species of birds hit in each strike. We hypothesize that bird species are more likely to be hit during their migration season, and that flocks of birds are also more likely to be hit during migration. As more than 350 North American bird species migrate, we also hypothesize that there will be a significant increase in bird strikes in general in the migration months.

So we want the months 3 and 4, and months 8, 9, 10 as the migration months.

### Plot strike frequency by migration season

```{r cars}
snarge <- read.csv("~/Desktop/Snarge/database.csv")

#make chart of number of hits per month
hitsbymonth <- snarge %>%
  select(c(Incident.Month, Species.Name)) 

#create a season catagory
hitsbymonth$migrationseason <- "yes"
hitsbymonth$migrationseason <- ifelse(hitsbymonth$Incident.Month %in% (3:4) | hitsbymonth$Incident.Month %in% (8:10), hitsbymonth$migrationseason, "no")

#change month names
hitsbymonth$Incident.Month <- as.factor(hitsbymonth$Incident.Month)
levels(hitsbymonth$Incident.Month) <- c("Jan", "Feb", "Mar","Apr", "May", "Jun","Jul", "Aug", "Sep","Oct", "Nov", "Dec")

#create an 'incident count' catagory
hitsbymonth$count <- 1

#plot
ggplot(data=hitsbymonth, aes(x = Incident.Month, y= count, fill = migrationseason)) +
    xlab("Month") +
    ylab("Incident Count") +
    geom_bar(stat="identity") +
  theme_classic() +
  scale_fill_manual(values=c("#999999", "#E69F00"), 
                       name="Migration\nSeason",
                       labels=c("Off months", "On months")) +
    theme(axis.text.x=element_text(angle=60,vjust=0.5))

```

## Plot strike severity by month

```{r pressure}

#plot SIZE of impact
quant <- snarge %>% 
  select(c(Incident.Year, Incident.Month, Incident.Day, Species.ID, Species.Name, Species.Quantity))

#levels(as.factor(quant$Species.Quantity))
#               1   10-Feb   11-100 Over 100 
#    4477   146563    21852     1166       46

#replace each range with the middle value of that range
quant$Species.Quantity <- gsub("10-Feb", "5", as.character(quant$Species.Quantity))
quant$Species.Quantity <- gsub("11-100", "50", as.character(quant$Species.Quantity))
quant$Species.Quantity <- gsub("Over 100", "100", as.character(quant$Species.Quantity))

#change to integer class
quant$Species.Quantity <- as.integer(quant$Species.Quantity)

#put quantity data into hitsbymonth dataframe
hitsbymonth$quant <- quant$Species.Quantity

#remove NAs and count total hits per month
hitsbymonth %>% 
  group_by(Incident.Month) %>% 
  na.omit() %>% 
  summarise(sum=sum(quant))

ggplot(data=hitsbymonth, aes(x = Incident.Month, y=quant, fill = migrationseason)) +
    xlab("Month") +
    ylab("Incident Count") +
    geom_bar(stat="identity") +
  theme_classic() +
  scale_fill_manual(values=c("#999999", "#E69F00"), 
                       name="Migration\nSeason",
                       labels=c("Off months", "On months")) +
    theme(axis.text.x=element_text(angle=60,vjust=0.5))

##make a stacked barchart that has bigger flocks colour coded as darker blue
hitsbymonth$quant <- as.factor(hitsbymonth$quant)

ggplot(data=hitsbymonth, aes(x = Incident.Month, y=count, fill = quant)) +
    xlab("Month") +
    ylab("Incident Count") +
    geom_bar(stat="identity") +
  theme_classic() +
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9", "#FF5733"), 
                       name="Migration\nSeason",
                       labels=c("1", "1-10", "10-100", "100+")) +
    theme(axis.text.x=element_text(angle=60,vjust=0.5))

#try filtering out single hits...

big <- hitsbymonth %>% filter(quant == 5 | quant == 50 | quant == 100)

ggplot(data=big, aes(x = Incident.Month, y=count, fill = quant)) +
    xlab("Month") +
    ylab("Incident Count") +
    geom_bar(stat="identity") +
  theme_classic() +
  scale_fill_manual(values=c("#BFF6FB", "#24C4D0", "#086A72"), 
                       name="Flock\nSize",
                       labels=c("2-10", "10-100", "100+")) +
    theme(axis.text.x=element_text(angle=60,vjust=0.5))

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{Orhun stuff}
#Setup
library(dplyr)
library(ggplot2)
Data <- snarge

# There are 715 unique labels for species for the data not counting blank cells
unique(Data$Species.Name)

#remove unneeded variables
Data <- Data %>% 
  select(c(Incident.Year, Incident.Month, Incident.Day, Species.ID, Species.Name, Species.Quantity))

#break up data into 5 year chunks
Data$Chunk <- Data$Incident.Year
Data$Chunk <- gsub("1990|1991|1992|1993|1994", "1", Data$Chunk)
Data$Chunk <- gsub("1995|1996|1997|1998|1999", "2", Data$Chunk)
Data$Chunk <- gsub("2000|2001|2002|2003|2004", "3", Data$Chunk)
Data$Chunk <- gsub("2005|2006|2007|2008|2009", "4", Data$Chunk)
Data$Chunk <- gsub("2010|2011|2012|2013|2014|2015", "5", Data$Chunk)

#Creating table arranged by frequency
X <- table(Data$Species.Name)
X <- sort(table(Data$Species.Name),decreasing=T)

#Removing Unknown species and renaming column
X = as.data.frame(X)
X = X[- grep("UNKNOWN", X$Var1),]
names(X)[names(X) == "Var1"] = "Species"

#Narrow to top 10 since its not possible or organized to display all species
X1 = X[1:10,]

#Plot Setup
theme_pubworthy = function (base_size = 12, base_family = "") {
  theme_classic(base_size = base_size, base_family = base_family) %+replace% 
    theme(
      axis.text = element_text(colour = "black"),
      axis.title.x = element_text(size=18),
      axis.text.x = element_text(size=6),
      axis.title.y = element_text(size=18,angle=90),
      axis.text.y = element_text(size=10),
      axis.ticks = element_blank(), 
      panel.background = element_rect(fill="white"),
      panel.border = element_blank(),
      plot.title=element_text(face="bold", size=24),
      legend.position="none"
    ) 
}

#make a line for each different species, and have their hits  over all the years
#need to make a dataframe with each of top 10 species as a column, with their frequency for each year

#this had the problem of grabbing many types of gulls/sparrows, fixed now
topten <- Data %>% filter(grepl("MOURNING DOVE|^GULL$|KILLDEER|AMERICAN KESTREL|BARN SWALLOW|HORNED LARK|EUROPEAN STARLING|^SPARROW$|ROCK PIGEON|RED-TAILED HAWK", Species.Name))
#summary(as.factor(as.character(topten$Species.Name)))

topten <- topten %>% 
  group_by(Incident.Year, Species.Name) %>% 
  tally()

topten$Species <- topten$Species.Name

#plot the top ten
ggplot(topten, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
    theme_classic() +
  xlab("Year") +
    ylab("Strike Count") +
  scale_fill_discrete(name="Species")

#ten to twenty
X2 = X[11:20,]

toptwen <- Data %>% filter(grepl("EASTERN MEADOWLARK|^HAWK$|CANADA GOOSE|RING-BILLED GULL|CLIFF SWALLOW|^BLACKBIRD$|HERRING GULL|^SWALLOW$|BARN OWL|WESTERN MEADOWLARK", Species.Name))
#summary(as.factor(as.character(topten$Species.Name)))

toptwen <- toptwen %>% 
  group_by(Incident.Year, Species.Name) %>% 
  tally()

toptwen$Species <- toptwen$Species.Name

#plot the top twen
ggplot(toptwen, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
    theme_classic() +
  xlab("Year") +
    ylab("Strike Count") +
  scale_fill_discrete(name="Species")

#try all 20 together
toptwenty <- Data %>% filter(grepl("EASTERN MEADOWLARK|^HAWK$|CANADA GOOSE|RING-BILLED GULL|CLIFF SWALLOW|^BLACKBIRD$|HERRING GULL|^SWALLOW$|BARN OWL|WESTERN MEADOWLARK|MOURNING DOVE|^GULL$|KILLDEER|AMERICAN KESTREL|BARN SWALLOW|HORNED LARK|EUROPEAN STARLING|^SPARROW$|ROCK PIGEON|RED-TAILED HAWK", Species.Name))

toptwenty <- toptwenty %>% 
  group_by(Incident.Year, Species.Name) %>% 
  tally()

toptwenty$Species <- toptwenty$Species.Name

#plot top twenty
ggplot(toptwenty, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
    theme_classic() +
  xlab("Year") +
    ylab("Strike Count") +
  scale_fill_discrete(name="Species")

##For blackbird
blackbird <- Data %>% filter(grepl("BLACKBIRD", Species.Name))
blackbird <- blackbird %>% 
  group_by(Incident.Year, Species.Name) %>% 
  tally()
blackbird$Species <- blackbird$Species.Name
ggplot(blackbird, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
    theme_classic() +
  xlab("Year") +
    ylab("Strike Count") +
  scale_fill_discrete(name="Species", position = "top")

#For swallow
swallow <- Data %>% filter(grepl("SWALLOW", Species.Name))
swallow <- swallow %>% 
  group_by(Incident.Year, Species.Name) %>% 
  tally()
swallow$Species <- swallow$Species.Name
ggplot(swallow, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
    theme_classic() +
  xlab("Year") +
    ylab("Strike Count") +
  scale_fill_discrete(name="Species")

#For sparrow
sparrow <- Data %>% filter(grepl("SPARROW", Species.Name))
sparrow <- sparrow %>% 
  group_by(Incident.Year, Species.Name) %>% 
  tally()
sparrow$Species <- sparrow$Species.Name
ggplot(sparrow, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
    theme_classic() +
  xlab("Year") +
    ylab("Strike Count") +
  scale_fill_discrete(name="Species")

#For gull
gull <- Data %>% filter(grepl("GULL", Species.Name))
gull <- gull %>% 
  group_by(Incident.Year, Species.Name) %>% 
  tally()
gull$Species <- gull$Species.Name
ggplot(gull, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
    theme_classic() +
  xlab("Year") +
    ylab("Strike Count") +
  scale_fill_discrete(name="Species")

#for dove
dove <- Data %>% filter(grepl("DOVE", Species.Name))
dove <- dove %>% 
  group_by(Incident.Year, Species.Name) %>% 
  tally()
dove$Species <- dove$Species.Name
#nice dove
ggplot(dove, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
  theme(legend.position = "bottom", legend.title = element_blank(),
  panel.border = element_blank(),  
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "grey"),
  legend.key=element_blank()
  ) +
  xlab("Year") +
  ylab("Strike Count") +
  guides(fill=guide_legend(nrow=2,byrow=TRUE))


#nicer blackbird and gull
ggplot(blackbird, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
  theme(legend.position = "bottom", legend.title = element_blank(),
  panel.border = element_blank(),  
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "grey"),
  legend.key=element_blank()
  ) +
  xlab("Year") +
  ylab("Strike Count") +
  guides(fill=guide_legend(nrow=4,byrow=TRUE))

   theme_set(theme_bw()) +

ggplot(gull, aes(x= Incident.Year, y=n, colour= Species)) + 
  geom_line(aes(colour = Species)) +
   geom_point(aes(colour = Species)) + 
  theme(legend.position = "bottom", legend.title = element_blank(),
  panel.border = element_blank(),  
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "grey"),
  legend.key=element_blank(),
  ) +
  xlab("Year") +
  ylab("Strike Count")